---
title: "PM566-lab4"
author: "Weixi Pan"
format:
  html:
    embed-resources: true
---

Step1
```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

Step2
```{r}
library(R.utils)
library(lubridate)
#Remove temperatures less than -17C
met <- met[met$temp > -17][elev == 9999.0, elev := NA]
#Generate a date variable using the functions as.Date()
met$date <- as.Date(paste(met$year, met$month, met$day, sep = "-"))
#Using the data.table::week function, keep the observations of the first week of the month.
met[,table(week(date))]
met <- met[week(date)==31]
```
```{r}
library(dplyr)
library(tidyverse)
#Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met_avg <- met[,.(
  temp     = mean(temp,na.rm=TRUE),
  rh       = mean(rh,na.rm=TRUE),
  wind.sp  = mean(wind.sp,na.rm=TRUE),
  vis.dist = mean(vis.dist,na.rm=TRUE),
  dew.point= mean(dew.point,na.rm=TRUE),
  lat      = mean(lat),
  lon      = mean(lon), 
  elev     = mean(elev,na.rm=TRUE)
), by=c("USAFID")]
```
```{r}
library(data.table)
#Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
met_avg[,region := fifelse(lon >= -98 & lat > 39.71,"NE",
              fifelse(lon < -98 & lat > 39.71,"NW",
              fifelse(lon < -98 & lat <= 39.71,"SW", "SE")))]

table(met_avg$region)
```
```{r}
#Create a categorical variable for elevation as in the lecture slides
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```

step3
```{r}
library(ggplot2)
#Try using geom_violin instead
met_avg[!is.na(elev_cat)] %>% 
  ggplot()+
  geom_violin(mapping=aes(x=elev_cat, y=temp, fill=elev_cat))
  facet_wrap(~elev_cat, nrow=1)
```
The result shows that the median temperature in high elev is lower than the low elev.

Step4
```{r}
#Use geom_jitter with stat_smooth to examine the association between dew point and wind speed by region
met_avg[!is.na(region)]%>%
  ggplot()+
  geom_jitter(mapping=aes(x=dew.point, y=wind.sp, fill=region))+
  geom_smooth(mapping = aes(x = dew.point, y = wind.sp, linetype = region))
```
Except the north east area, wind speed of most areas stays same when dew point change.

step5
```{r}
#Use geom_bar to create barplots of the weather stations by elevation category colored by region
met_avg[!is.na(elev_cat) & elev_cat != "clear"] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = elev_cat, fill = region), position = "dodge")+
   scale_fill_brewer(palette= "Blues")+
   labs( x= "Elevation Category", y= "Number of Stations", title = "Weather Station")+
  theme_bw(base_family = "Times")
```
In high elevation area, the weather station build mostly in northeast place (around 1250), and northwest place holds the least (no over 750). In low elevation area, the weather station build mostly in southeast place (around 2500), and northwest place holds the least (no over 100). 

Step6
```{r}
#Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

#1.dew point
met_avg[!is.na(dew.point)] %>%
  ggplot() + 
    stat_summary( mapping = aes(y = dew.point, x = region), fun.data="mean_sdl", geom = "errorbar")
```
Dew point is wilder range in northwest and southwest area.
```{r}
#2.wind speed
met_avg[!is.na(wind.sp)] %>%
  ggplot() + 
    stat_summary( mapping = aes(y = wind.sp, x = region), fun.data="mean_sdl", geom = "errorbar")
```
Wind speed is almost same in 4 regions.

step7
```{r}
#Make a map showing the spatial trend in relative humidity in the US
library(leaflet)
met_avg <- met_avg[!is.na(rh)]

# Generating a color palette
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal

top10rh <-  met_avg[rank(-rh)<=10]

rhmap <- leaflet(met_avg) %>%  
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humidity', opacity=1)
rhmap <- addMarkers(rhmap, data=top10rh, lng=~lon, lat=~lat,label = ~rh)
rhmap
```
The humidity increases from west to east in the US. And a number of high humidity cities are located in the north place.

Step8
```{r}
# Use a ggplot extension
library(ggforce)
#> Loading required package: ggplot2
ggplot(met_avg, aes(rh, temp, colour = region)) +
  geom_point() +
  facet_zoom(x = region == "versicolor")

```

The `echo: false` option disables the printing of code (only output is displayed).
